---
title: "Matrices"
output: html_document
---

# Matrices

A matrix is a two dimensional data structure that is similar to a vector but additionally contains the dimension attribute.

## 1. Creating

When creating a matrix, the last two arguments to matrix tell it the number of rows and columns the matrix should have.

#### 1.1 Matrix Code Example

You can use the byrow=TRUE argument to create a matrix by rows instead of by columns as shown below:

```{r}
## Example 
# ---
# Question: Let's create a matrix mymat
# ---
# 
mymat <- matrix(1:12,4,3)

# And then print out mymat
# ---
mymat
```
#### 1.2 Matrix Code Example

```{r}
## Example 
# ---
# Question: Let's use the byrow=TRUE argument to create a matrix 
# by rows instead of by columns as shown below
# ---
mymat <- matrix(1:12,ncol=3,byrow=TRUE)

# And then print out the mymat variable below
mymat
```
## 2. Naming

#### 2.1 Matrix Naming Code Example

```{r}
## Example 
# ---
# In order to remember what is stored in a matrix, you can add the names of the columns and rows. 
# This will also help you to read the data as well as select elements from the matrix.
# ---
# Question: Lets create the vectors kenya, ethiopia and chad
# ---
# OUR CODE GOES BELOW
# 
kenya <- c(460.998, 314.4) 
ethiopia <- c(290.475, 247.900) 
chad <- c(309.306, 165.8)

# then create a matrix geography_matrix
# ---
# 
geography_matrix <- matrix(c(kenya, ethiopia, chad), nrow = 3, byrow = TRUE)

# Uncomment the following two vectors; location and countries which will be used for naming
# ---
#
location <- c("Lat", "Long")
countries <- c("Kenya", "Ethiopia", "Chad")

# Then also uncomment the line below and so as to name the columns with location
 
colnames(geography_matrix) <- location

# More on uncomment also the following line so as to name the rows with countries

rownames(geography_matrix) <- countries

# And lastly print out geography_matrix

geography_matrix
```
```{r}
## Challenge 
# ---
# Question: Create a matrix family with column names Name, Age, Gender and Occupation. 
# Populate it with 5 your own family members.
# ---
Name <- c("John","James","Luke","Mark","Bathlomew")
Age <- c(34,35,36,37,38)
Gender <- c("M","M","M","M","M")
Occupation <- c("Missionary","Missionary","Missionary","Missionary","Missionary")

family <- matrix(c(Name,Age,Gender,Occupation),nrow = 5, byrow = FALSE)
colnames(family) <- c("Name","Age","Gender","Occupation")
family
```
## 3. Adding a Column

You can add a row to a matrix using the rbind() function.

#### 3.1 Adding a Column Code Example

```{r}
## Example 
# ---
# Question: Create the matrix x below
# ---
# OUR CODE GOES BELOW
# 
x <- matrix(1:9, nrow = 3)

# Then add a column as shown

cbind(x, c(1, 2, 3))
```
```{r}
## Challenge 
# ---
# Question: Add a column residence to your fictional family matrix that you had created earlier
# ---
# OUR CODE GOES BELOW
cbind(family, c("Jerusalem","Jerusalem","Jerusalem","Jerusalem","Jerusalem"))
```

## 4. Adding a Row

#### 4.1 Adding a Row Code Example
```{r}
## Example 
# ---
# Question: Creating a matrix x
# ---
# 
x <- matrix(1:9, nrow = 3)

# You can also add a row using the cbind() function
# ---
# 
rbind(x,c(1,2,3))
```


```{r}
## Challenge 
# ---
# Question: Add a fictional character to your fictional family matrix
rbind(family,c("John Mark", 26, "M", "Missionary"))
```

## 5. Selecting a Matrix

To select an element of a matrix, one needs to specify both the row and the column as shown:


```{r}
## Example 
# ---
# Question: Select the following matrix
# ---
# OUR CODE GOES BELOW
# 
x <- matrix(1:9, nrow = 3)

# Select the elements from the above matrix 
# By uncommenting the following lines.
# ---
# 
x[1,3] # select the element at 1nd row, 3rd column
x[2, ] # the 2nd row
x[ ,3] # the 3rd column
```


```{r}
## Challenge 
# ---
# Question: Select the last member of your family member 
# ---
family[5,]
```


```{r}
## Challenge 
# ---
# Question: Select the first member of your family member 
# ---
family[1,]
```

## 6. Operations


### Matrix addition &amp; subtraction

#### 6.1 Matrix Addition &amp; Subtraction Code Example


```{r}
## Example 
# ---
# Question: Matrix addition and subtract require the matrices to have the same dimensions. 
# Let's start by creating matrices x and y.
# ---
# OUR CODE GOES BELOW
# 
x <- matrix(c(3, 9, -1, 4, 2, 6), nrow = 2)
y <- matrix(c(5, 2, 0, 9, 3, 4), nrow = 2)

# Print matrix x below
x
# Print the matrix y below
y
# Add the matrices and print out the result
x + y

# Subtract the matrices
x - y
```

### Matrix Multiplication &amp; Division

#### 6.2 Matrix Multiplication &amp; Division Code Example


```{r}
## Challenge
# ---
# Question: Let's create two 2 x 3 matrices x and y
# ---
# OUR CODE GOES BELOW
# 
x <- matrix(c(3, 9, -1, 4, 2, 6), nrow = 2)
y <- matrix(c(5, 2, 0, 9, 3, 4), nrow = 2)

# And then multiply these matrices, assign the result to to the variable z
z <- x * y

# Now print out the matrix z
z
# Lastly divide matrix x by y and assign the result to the variable z
z <- x/y
z
```